Covariates analysis of MSBB reprocessed RNASeq data

### 2016-06-02 code/Rmd/MSSM_RNASeq_Reprocessed.Rmd
Minor modifications

# Sex
vcd::mosaic(~ SEX + Diagnosis + BrainRegion, data = METADATA)
```
### Quality Control
```{r check.sex}
# Query biomart for Y chromosome genes
ensembl = useMart('ensembl', dataset = 'hsapiens_gene_ensembl', host = 'www.ensembl.org')
EnsembleInY <- getBM(attributes = c("ensembl_gene_id"),
                       filters = "chromosome_name", 
                       values = 'Y',
                       mart = ensembl)

@@

tmp1 = t(COUNT[YChromGenes$Gene.ID,]) %>%
  rowSums %>% rownameToFirstColumn('SampleID') %>%
  dplyr::rename(CPM.Sum = DF) %>%
  left_join(COVARIATES %>% 
              rownameToFirstColumn('SampleID'))

ggplot(tmp1, aes(x = SEX, y= log2(CPM.Sum))) + geom_boxplot()

@@

tmp1 = t(COUNT[XChromGenes$Gene.ID,]) %>%
  rowSums %>% rownameToFirstColumn('SampleID') %>%
  dplyr::rename(CPM.Sum = DF) %>%
  left_join(COVARIATES %>% 
              rownameToFirstColumn('SampleID'))

ggplot(tmp1, aes(x = SEX, y= log2(CPM.Sum))) + geom_boxplot()
```
### CPM Normalisation
Preprocess counts matrix and metadata. Determine design matrix for normalisation and differential expression analysis.

>>>>

# Sex
vcd::mosaic( Diagnosis + SEX ~ BrainRegion, data = METADATA)
```
### Quality Control
```{r check.sex}
# Query biomart for Y chromosome genes
ensembl = useMart('ensembl', dataset = 'hsapiens_gene_ensembl', host = 'www.ensembl.org')
EnsembleInY <- getBM(attributes = c("ensembl_gene_id"),
                       filters = "chromosome_name", 
                       values = 'Y',
                       mart = ensembl)

@@

PROCESSED_COUNTS = COUNT
PROCESSED_COUNTS = PROCESSED_COUNTS[rowSums(COUNT) != 0,]

VOOM.GENE_EXPRESSION = voom(PROCESSED_COUNTS, design=NULL, plot=F)

tmp1 = apply(VOOM.GENE_EXPRESSION$E[intersect(rownames(VOOM.GENE_EXPRESSION$E), YChromGenes$Gene.ID),], 2, mean) %>%
  rownameToFirstColumn('SampleID') %>%
  dplyr::rename(mean.log2.CPM = DF) %>%
  left_join(COVARIATES %>% 
              rownameToFirstColumn('SampleID'))

pl = list()
pl[[1]] = ggplot(tmp1, aes(x = SEX, y= mean.log2.CPM)) + geom_boxplot() + ggtitle('Mean expression of Y chromosome genes')

@@

tmp1 = apply(VOOM.GENE_EXPRESSION$E[intersect(rownames(VOOM.GENE_EXPRESSION$E), XChromGenes$Gene.ID),], 2, mean) %>%
  rownameToFirstColumn('SampleID') %>%
  dplyr::rename(mean.log2.CPM = DF) %>%
  left_join(COVARIATES %>% 
              rownameToFirstColumn('SampleID'))

pl[[2]] = ggplot(tmp1, aes(x = SEX, y= mean.log2.CPM)) + geom_boxplot() + ggtitle('Mean expression of X chromosome genes')

multiplot(plotlist = pl, cols = 2)
```
### CPM Normalisation
Preprocess counts matrix and metadata. Determine design matrix for normalisation and differential expression analysis.
